Network Origins of Social Distinction

### 2025-07-05 manuscript.tex

\documentclass[12pt]{article}
\usepackage[left=1.15in, right=1.15in, top=1in, bottom=1.5in]{geometry}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\bibpunct[: ]{(}{)}{;}{a}{}{,} 
\usepackage{authblk}
\renewcommand\Authfont{\normalsize}
\renewcommand\Affilfont{\footnotesize}

@@

\newpage
\begin{abstract}	
In this paper a new network measure, distinction centrality, is introduced. The measure is derived from Bourdieu's theory of the competition for distinction in a field where actors act both with agency and within the constraints imposed by other actors. The measure subtracts the autonomy of each actor \textit{i} within the network from their constraint. Autonomy is measured using eigenvector centrality for each actor \textit{i}. In contrast, constraint is measured by calculating the average eigenvector of nodes \textit{i}'s neighbors in a recalculated network with node \textit{i} removed. A scalar is then applied to create mathematical and sociological equivalence between autonomy and constraint. The resulting node-level measures create a measure of distinction, where the uniqueness of a node's associations is more heavily rewarded than under eigenvector centrality. The resulting measure is applied to both Padgett and Ansell's \citeyearpar{padgett1993robust} Medici Network and Zachary's \citeyearpar{zachary1977information} Karate Club to show that the observed actions of the agents involved are consistent with distinction maximization. The resulting measure can be used to both forecast network transformations and anticipate competitive changes resulting from external network shocks.    
\end{abstract}

\newpage
\section{Distinction Beyond Taste}
Popular within cultural sociology, the idea of distinction has generated a rich literature to explain cultural differentiation. Distinction, as the expression of status and uniqueness, is a social end in itself. The question of cultural distinction is often treated empirically as a matter of the deference given to symbolic forms of differentiation \citep{bourdieu1984distinction}. Broader social distinction is often measured as the shadow of cultural distinction or collapsed into the concept of status more generally, which ignores questions of how network constraint impacts the capacity of actors to be distinct. A generalized, measured ordering of social distinction that is derivative from a network of concrete social relations would give Bourdieu's generalized conception of distinction a broader methodological base from which to conduct empirical analysis.

\subsection{What is Distinction?}
Pierre Bourdieu was interested in the competition among people for positions within the hierarchy. Bourdieu defined the field as a social space where actors operate within socially negotiated boundaries that arbitrate relevant competitive claims over position. In interplay with the field, individuals develop a habitus of embodied dispositions, expressed as a cognitive ease with particular types of settings, objects, and performances. An actor's position in the field influences their habitus. In the same way, their habitus determines their future possibilities within the field, while also structuring the opportunities available for others in the field. Actors compete for positions in the field via various forms of capital, or resources that are socially-agreed upon as desirable within a field, thereby facilitating the formation of connections that both order and reflect the hierarchy of the field. The interaction of the desirability and uniqueness of the capitals possessed by an individual determines their place in the field's positional hierarchy.

Bourdieu \citeyearpar{bourdieu1984distinction} most famously used the conceptualization of distinction to explain the persistence of the relationship between class and cultural hierarchies. Cultural elites in France distinguished themselves from their inferiors by adopting differentiated cultural practices that had undergone a process of consecration, which set them apart from the common. There is a strong correlation between those in the position of cultural and economic elites. Still, such correlation is not guaranteed at the individual level even if it is relevant at the level of the fields.

While largely a concept used by those who research cultural practices, the pursuit of distinction is a general expression of the interaction between status and uniqueness within a field of positions \citep{anheier1995forms}. For example, while cultural practices may be a marker of difference, the absence of association with others outside the milieu, distinction is also a matter of agency within the milieu, a question of the constraint imposed by the set of relationships within which the actor is embedded. For Bourdieu, the struggle for position in a field was a product of a field of action and a field of forces \citep{martin2003field}. The field of action describes the agency of actors actively advancing their position. In contrast, the field of forces describes the counteracting pressures faced by the other participants whose interests would be upset by positional transformations.

In relation to theory, network status is another name for \textit{influence} \citep{friedkin1991theoretical}, and uniqueness for \textit{autonomy} in a social environment \citep{burt1980autonomy}. Translated to network terms, Bourdieu's field of action can be understood as the status hierarchy of the network, while the field of struggle is reflective of the constraint neighbors put on an actor. An actor with high influence is well-connected to others who can be mobilized for effective action. An actor with high autonomy is not constrained by their pattern connections or forced into action by them \citep{burt76}. We often use context-specific words to denote position in the hierarchy and ability to act within that hierarchy. Nevertheless, we use status and autonomy below as general terms to describe the ability of actors to take action versus the level of imposition from the forces of constraint.

In developing a broader understanding and measure of social distinction, actor-level incentives and structural impediments to action become apparent. When we understand the underlying motivation actors have to proactively reshape their network connections so as to achieve more distinct positions, historical patterns of network change can be better accounted for and sometimes even properly forecaste.

\subsection{Position as the Relational Basis for a Distinction Measure}
Bourdieu's separation of distinction into habitus and position is no sociological problem at all, but creates the analytical problem of how an actor's network position creates dispositions and how dispositions create the opportunity for network mobility. Both dispositions and positions are fundamentally relational outcomes in sociological terms, but in creating a network measure of distinction, only the relational basis of position can be considered.

However, to the extent that position and habitus reflect each other through their symbiotic fabrication, the ability to empirically capture only one of the two components is sufficiently accurate to measure a social system's distinction hierarchy with confidence over a short window of time. Actors simply should not have a network position that does not match their habitus, and their ability to radically change positions is governed by their current habitus and social position. If the transformation of habitus is a stepwise interplay with positional movement that unfolds over time, the measure becomes less informative if capturing networks over extended time periods. However, even over long periods of time this theoretical collapse that permits a network measured methodological shortcut may still be adequate in a moderately stable social environment.

To put the measure of distinction on a firm theoretical foundation, it is first important to make explicit the relationship between empirical data and social theory that is the basis of social network analysis. Network measures are fundamentally based on the idea that researchers are in some way capturing a relationship between two objects. A tie between two nodes is capturing in some way their connection however so defined. Such connections can be broadly classified as mutual in the case of the relationship quality being equal in both directions, or directional in the case of the qualities being unequal. In the absence of direct measurement of qualities however, many networks are measured as consisting of mutual ties. However, even in the case of networks composed of only mutual ties, hierarchy can still be assessed from the pattern of relationships. In such networks, we are still able to rank the direction of power and deference through the absence of ties between lower and higher nodes. For example, in an organizational pyramid, we would be able to detect the CEO from the presence of ties with the vice presidents and the absence of ties with middle management and production workers. We can do this through the measurement of a node's network status.

The first aspect of social distinction generally is \textit{status}. Status is the internal stratification of all members of the network, along with connections of deference and obligation. In the case of politics, status is the ability to access those who themselves have higher levels of access, with those most central to the access network having the ability to make decisions. This can be captured traditionally by measures of status such as eigenvector centrality \citep{bonacich1972norms}. Status is higher for actors when the people they are connected to are themselves important, and accessibility is just a political reframing of the context. The internal stratification of individuals is determined by who has greater connections to those who have greater connections, with those with few connections to other unimportant people being at the bottom of the hierarchy \citep{bonacich1972norms}.

@@

The interaction of influence and autonomy combine to create an actor’s overall distinction in the network. The combination of influence and autonomy results in capacity, and the social reward for this capacity is distinction. An increase in one aspect can often, but not necessarily, lead to a decrease in the other. For example, a connection with a someone with high accessibility other can lead to a decrease in one’s autonomy, and new access comes with new constraints. However, the trade-off can maximize one’s overall distinction, so there are gains from the connection. Without such gains, the connection would not be formed.

\section{A Generalized Measure of Network Distinction}
Below we present a measure of distinction centrality, determine the appropriate application of a scalar factor, and point to the availability of an R program so that researcher's may easily implement the measure in their own research.

@@

A formalization of this interaction between status and uniqueness allows for a measure of network distinction. Distinction as a theoretical concept is composed of status mediated by autonomy. A distinction measure can reflect this by considering an actor'ss status and then discounting their status by the status of their neighbors independent of the actor. The more reliant an actors neighbors are on the actor for their status, the lower that actors ability to exert their own influence, and thus the lower their distinction. In other words, autonomy is determined by how much one's neighbors need you to sustain their own position in the network.

Distinction centrality can be measured through the implementation of the following algorithm.
\[\beta_i=Sx_i-{\frac{1}{n_{j=i}}\sum_{{j}\in{A},i\neq{j}}x_{j}} \]
Let \(\beta\) be the distinction centrality of a given individual \textit{i}. The first term captures the status of a given actor. Let \textit{x} be the eigenvector centrality \citep{bonacich72} of the designated individual. Let S be a scalar that adjusts the social value placed on status compared to constraint, to be discussed below. The second terms captures the element of autonomy. Removing individual \textit{i} from the network, compute the eigenvectors of the remaining individuals and sum those that are connected to individual \textit{i}, taking the average eigenvector of neighbors \textit{j} as the constraint confronting node \textit{i} in the network. This is substantively the same mechanism as conducted in a jackknife resampling procedure, except rather than forming a sampling distribution, we are measuring local constraint. Finally, subtract the status of the average neighbor connected to \textit{i} from \textit{i}’s scalar adjusted eigenvector centrality in the whole network.

If a node is an isolate, its eigenvector centrality should be set at 0. In the case of a network with no connections (such as a star graph), it can accidentally be calculated as 1.00 in programs, but given that it is still a network of no connections, setting isolates at 0.00 is theoretically appropriate when considering that status is a reflection of connections. It would be inappropriate for an isolate to have the same status as the highest status node in a large network. Additionally, given the second step of subtraction, it is only theoretically meaningful if set as 0 in the case of distinction centrality.

\subsection{Scalar Adjustment}
First, a scalar is required because relative value of status and autonomy is not inherently fixed in a social system. We might imagine situations where one desires more status at the expense of autonomy if the rewards to status are higher, or value autonomy more highly if the rewards to status are low. There cannot be an assumed unadjusted equivalence between the resulting eigenvectors of the two terms for this theoretical level.

Additionally, because of the hierarchical nature of networks, the average constraint in the network is larger than the average status. The high average constraint of their neighbors is often much larger than the impact of status. The result is that, at the level of the entire network, that those who are deeply embedded among the most elites, without their own basis of supporters, are very indistinct when considering the high constraints placed on them by their high status alters. Thus, the calculated results without a scalar can fail to capture the appropriate value on status as compared to constraint.

While an arbitrary scalar is possible, it can lead to erroneous results when tested against empirical networks. However, a scalar can be derived from the network itself because of the field positional interplay of status and constraint. Imagining two actors of equal status and personal constraint, one in an averagely more constrained system and the other in a less averagely constrained system, with average status otherwise equal. In such a case, the status of an actor in the more constrained system has a greater impact on their social distinction than in the less constrained system. With high constraint, status takes on greater importance in determining social distinction because it compels deference. In other words, it is the difference between hierarchy with constraint, and hierarchy without constraint. Status with weak constraint means a weaker compulsion of deference. The structure of constraint redounds into the impact of status on social distinction beyond its own contribution.

With this social understanding of how constraint interacts with the transmission of status, an endogenous scalar can be derived from the network. The average constraint of actors within in the network divided by the average status of actors in the network creates a scalar that reflects the relationship between status and constraint.

@@

Under conditions of high average constraint and low average status, the scalar is large, reflecting the importance of status in determining social distinction. Under conditions where average constraint is low and average status is high, the scalar is weaker, reflecting a system where status is less important in determining social distinction. Under very large-scale conditions, the averages of status and constraint should trend towards 1, weakening the necessity of the scalar. However, under constrained conditions, particularly elite networks with missing data of the social scaffolding that created the network elites, the scalar is more important for explaining action because the range of eigenvector network measures are otherwise being artificially expanded due to the missing data.

Without a scalar, networks with any degree of hierarchy see the average constraint swamp the average status because many low status nodes tie into high status nodes. The advantage of this scalar is to make the average scalar status in the network equal to the average constraint in the network. Thus, total status is equal to total constraint, and from there it becomes a matter of distribution among members. From there however, it allows for comparability between networks that would otherwise not be possible. With the understanding that we have, through the scalar, adjusted for a network where ones combination of status and constraint is better or worse, networks can be compared so long as the unobserved social context within the network changes are made is otherwise stable.

@@

\textit{Model (b)} presents an idealized depiction of a structural fold \citep{vedres2010structural}. The resulting measurement scores are in Table 2. While node 1 maintains high eigenvector centrality, the eigenvector centrality of the other nodes increases, while their constraint decreases. It is actually the constraint of Node 1 that increases, reaching 1.00 because all its neighbors are maximally interconnected in their own subgroups. This position of brokerage is more constrained than the star graph because the other actors are connected in subgroups. The actors in the subgroup are able to combine their interests to mitigate against the independence of Node 1 in the center. In an unadjusted measure of distinction, the different nodes become remarkably close in their distinction measure. With a scalar adjustment, the stratification increases, giving a bit more privilege to Node 1, but demonstrating the extent to which Node 1's distinction is eroded. In a maximally connected graph, the distinction of each node would also go to 0.00, as there is no network differentiation among nodes.

\textit{Model (c)} presents a graph that is 2-connected in the absence of node 1. Each subgraph is one connection short of forming complete subgraphs in the absence of Node 1. Compared to the structural fold,  connection 3-2 and connection 6-7 have been removed. Measurement results are in Table 3. Of the demonstrative networks here, this is the one that most approximates a three-tiered ordered hierarchy. The resulting scalar adjusted distinction measures well reflects this stratification, with Node 1 retaining its distinction, Node's 4 and 5 failing in the middle position, and Node's 2, 3, 6, and 7 falling to the bottom.

\textit{Model (d)} is the same 2-connected graph with the addition of a long tie connecting nodes 2 and 7. This long tie stratifies the network even more than in Model (c) bringing Nodes 2 and 7 quite close to 4 and 5. The formation of the long tie served to bolster the individual distinction of 2 and 7 more than the formation of the tie that would create the completely connected subgraphs as in Model (b). In fact, all nodes except 3 and 6 have higher distinction in such a case. This reflects the positional idea that distinction is a matter of stratification, and cohesion is at odds with creating stratification. While the status of Node's 2 and 7 actually increase beyond that of 4 and 5 due to the long tie, the constraint also increases as they are now both further constrained by the influence of Node 1 as their mutual neighbor.

\subsection{R Package}
The analyses below are conducted with a program that implements the above distinction centrality measure. Omar wrote it. He's a cool guy. Maybe he would like to do a whole separate article on it?

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/star.png}
            \caption{Star Graph}
            \label{fig:star}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/sf.png}
            \caption{Structural Fold}
            \label{fig:sf}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/intra.png}
            \caption{2-Connected Subgraphs}
            \label{fig:intra}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/inter.png}
            \caption{2-Connected Subgraphs with Long Tie}
            \label{fig:inter}
    \end{subfigure}
    \caption{}
    \label{fig:toys}
\end{figure}

@@

\section{Long Ties and Opportunities for Distinction}
Given that networks are not static, but evolving, the incentives to maximize distinction within a network can inform the evolution of culture as a function of network reorganization. As cultural capital is an outcome of both status and uniqueness, the same calculus of optimizing network position can be extended.

@@

\section{Capital Conversion as Network Integration of Fields}
One of the problems with the relationship between multiple fields is how capital might be converted between fields. The ``rate'' between fields has always been a matter of confusion because it is difficult to quantify. Yet, it is an essential feature of an individual's navigation between the positional hierarchies of various fields. Seeking to convert a distinct position from one field to a better position in another is a critical strategy in the overall positional portfolio of agents.

@@

\section{Limitations on Positional Agency}
One problem with the above measure is that is creates the spectre of limitless positional agency: if a connection is mutually distinction increasing, actors should pursue it. Given that distinction increasing positional changes are possible, but go un-enacted, there must be constraints to distinction maximizing action. Theoretical considerations must be made to account for costs that constrict connection formation for distinction optimization.

>>>>

\documentclass[12pt]{article}
\usepackage[left=1.15in, right=1.15in, top=1in, bottom=1.5in]{geometry}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[hidelinks]{hyperref}
\bibpunct[: ]{(}{)}{;}{a}{}{,} 
\usepackage{authblk}
\renewcommand\Authfont{\normalsize}
\renewcommand\Affilfont{\footnotesize}

@@

\newpage
\begin{abstract}	
This paper introduces the concept of \textit{distinction centrality}, developing a network metric designed to capture it. The metric is theoretically inspired by Bourdieu's theory of agonistic competition for distinction in social fields, where other actors and their actions provide opportunities and constraints to maximize social positioning. The basic idea behind the network metric of distinction is to subtract the \textit{autonomy} of each actor within the network from their \textit{constraint}, such that highly autonomous (and therefore less constrained) actors have more distinction. Autonomy is given by the focal actor's Eigenvector Centrality score (recording connections to well-placed others), and constraint is defined by the average Eigenvector Centrality score of each actor's neighbors in a hypothetical network where the focal actor is removed (indicating how others would benefit from that actor's exiting the network). A scalar is then applied to create mathematical and sociological equivalence between autonomy and constraint. The resulting node-level measures create a measure of \textit{distinction}, where the uniqueness of an actor's relations is more heavily rewarded than under the standard Eigenvector Centrality score. We analyze two classic datasets, Padgett and Ansell's \citeyearpar{padgett1993robust} Medici Network and Zachary's \citeyearpar{zachary1977information} Karate Club, to demonstrate that the observed actions of the agents involved are consistent with the maximization of distinction. We close by outlining how the resulting distinction measure can be used to both forecast network transformations and anticipate competitive changes resulting from external network shocks.    
\end{abstract}

\newpage
\section{Distinction Beyond Taste}
Popular within cultural sociology, the concept of distinction has generated a rich body of literature aimed at explaining cultural differentiation within social fields. From this perspective, the pursuit of distinction, as the expression of an actor's status and uniqueness, is an end in itself, providing the basic motivation for action \citep{martin2003field}. The question of cultural distinction is often treated empirically as a matter of the deference given to symbolic forms of differentiation enacted by designated institutional authorities \citep{bourdieu1984distinction}.

In this last respect, the more encompassing notion of \textit{social} distinction, implying an actor's embeddedness in a relational field \citep{erikson2013formalist-400, fuhse2021social-9cb}, is often measured as the shadow counterpart of cultural distinction or collapsed into the concept of status more generally \citep{podolny2010status-1a6}. This approach overlooks questions of how network constraints impact the capacity of actors to be distinct. A generalized, measured ordering of social distinction that is derivative from a network of concrete social relations would give Bourdieu's generalized conception of distinction a broader methodological base from which to conduct empirical analysis.

\subsection{What is Distinction?}
Bourdieu was primarily interested in the competition among actors for positions within the hierarchy implied by the field \citep{anheier1995forms, bourdieu1993field-8ad}. Bourdieu defined the field as a social space where actors operate within socially negotiated boundaries that arbitrate relevant competitive claims over position \citep{bourdieu1996rules-393}. In interplay with the field, individuals develop a habitus of embodied dispositions, expressed as cognitive ease with particular types of settings, objects, and performances. An actor's position in the field influences their habitus. Similarly, their habitus determines their future possibilities within the field, while also shaping the opportunities available to others in the field. Actors compete for positions in the field via various forms of capital or resources that are socially agreed upon as desirable within a field \citep{martin2003field}, thereby facilitating the formation of connections that both order and reflect the hierarchy of the field. The interaction of the desirability and uniqueness of the capitals possessed by an individual determines their place in the field's positional hierarchy.

Bourdieu \citeyearpar{bourdieu1984distinction} most famously used the conceptualization of distinction to explain the persistence of the relationship between class and cultural hierarchies. Cultural elites in France distinguished themselves from their inferiors by adopting differentiated cultural practices that had undergone a process of consecration, which set them apart from the common. There is a strong correlation between those in the position of cultural and economic elites. Still, such a correlation is not guaranteed at the individual level, even if it is relevant at the field level.

While largely a concept used by those who research cultural practices, the pursuit of distinction is a general expression of the interaction between status and uniqueness within a field of positions \citep{anheier1995forms}. For example, while cultural practices may be a marker of difference, the absence of association with others outside the milieu is also a matter of agency within the milieu, a question of the constraints imposed by the set of relationships within which the actor is embedded. For Bourdieu, the struggle for position in a field was a product of a field of action and a field of forces \citep{martin2003field}. The field of action refers to the agency of actors actively advancing their interests. In contrast, the field of forces describes the counteracting pressures faced by the other participants whose interests would be upset by positional transformations.

In relation to theory, network status is another name for \textit{influence} \citep{friedkin1991theoretical}, and uniqueness for \textit{autonomy} in a social environment \citep{burt1980autonomy}. Translated to network terms, Bourdieu's field of action can be understood as the status hierarchy of the network, while the field of struggle is reflective of the constraints neighbors put on an actor. An actor with high influence is well-connected to others who can be mobilized for effective action. An actor with high autonomy is not constrained by their pattern connections or forced into action by them \citep{burt76}. We often use context-specific words to denote position in the hierarchy and ability to act within that hierarchy. Nevertheless, we use ``status" and ``autonomy" below as general terms to describe the ability of actors to take action versus the level of imposition from the forces of constraint.

In developing a broader understanding and measure of social distinction, actor-level incentives and structural impediments to action become apparent. When we understand the underlying motivation actors have to proactively reshape their network connections so as to achieve more distinct positions, historical patterns of network change can be better accounted for and sometimes even properly forecasted.

\subsection{Position as the Relational Basis for a Distinction Measure}
Bourdieu's separation of distinction into habitus and position is not a sociological problem at all, but rather creates the analytical problem of how an actor's network position generates dispositions and how these dispositions create opportunities for network mobility. Both dispositions and positions are fundamentally relational outcomes in sociological terms; however, when creating a network measure of distinction, only the relational basis of position can be considered.

However, to the extent that position and habitus reflect each other through their symbiotic fabrication, the ability to empirically capture only one of the two components is sufficiently accurate to measure a social system's distinction hierarchy with confidence over a short window of time. Actors should not hold a network position that does not align with their habitus, and their current habitus and social position determine their ability to radically change positions. If the transformation of habitus is a stepwise interplay with positional movement that unfolds over time, the measure becomes less informative if it captures networks over extended periods. However, even over long periods, this theoretical collapse, which permits a network-measured methodological shortcut, may still be adequate in a moderately stable social environment.

To establish a firm theoretical foundation for the measure of distinction, it is first essential to make explicit the relationship between empirical data and social theory that forms the basis of social network analysis. Network measures are fundamentally based on the idea that researchers are capturing a relationship between two objects in some way. A tie between two nodes captures, in some way, their connection, however it is defined. Such connections can be broadly classified as mutual, in the case where the relationship quality is equal in both directions, or directional, in the case where the qualities are unequal. In the absence of direct measurement of qualities, however, many networks are measured as consisting of mutual ties. However, even in the case of networks composed of only mutual ties, hierarchy can still be assessed from the pattern of relationships. In such networks, we can still rank the direction of power and deference through the absence of ties between lower and higher nodes. For example, in an organizational pyramid, we can detect the CEO by observing the presence of ties with vice presidents and the absence of ties with middle management and production workers. We can do this through the measurement of a node's network status.

The first aspect of social distinction is \textit{status}. Status refers to the internal stratification of all network members, along with the connections of deference and obligation. In the case of politics, status refers to the ability to access those who themselves have higher levels of access, with those most central to the access network having the authority to make decisions. This can be captured traditionally by measures of status such as eigenvector centrality \citep{bonacich1972norms}. Status is higher for actors when the people they are connected to are themselves influential, and accessibility is merely a political reframe of the context. The internal stratification of individuals is determined by who has greater connections to those with even greater connections, with those having few connections to other important people being at the bottom of the hierarchy \citep{bonacich1972norms}.

@@

The interaction of influence and autonomy combines to create an actor’s overall distinction in the network. The combination of influence and autonomy results in capacity, and the social reward for this capacity is distinction. An increase in one aspect can often, but not necessarily, lead to a decrease in the other. For example, a connection with someone with high accessibility can lead to a decrease in one’s autonomy, and new access comes with new constraints. However, the trade-off can maximize one’s overall distinction, so there are gains from the connection. Without such gains, the connection would not be formed.

\section{A Generalized Measure of Network Distinction}
Below, we present a measure of distinction centrality, determine the appropriate application of a scalar factor, and indicate the availability of an R program to facilitate easy implementation of the measure in further wok.

@@

A formalization of this interaction between status and uniqueness allows for a measure of network distinction. Distinction as a theoretical concept is composed of status mediated by autonomy. A distinction measure can reflect this by considering an actor's status and then discounting their status by the status of their neighbors, independent of the actor. The more reliant an actor's neighbors are on the actor for their status, the lower the actor's ability to exert their influence, and thus the lower their distinction. In other words, autonomy is determined by how much one's neighbors rely on you to maintain their position in the network.

Distinction centrality can be measured using the following expression:
\begin{equation}
\beta_i=Sx_i-{\frac{1}{n_{j=i}}\sum_{{j}\in{A},i\neq{j}}x_{j}} 
\end{equation}

Let $\beta_i$ be the distinction centrality of a given individual \textit{i}. The first term captures the status of a given actor. Let \textit{x} be the eigenvector centrality \citep{bonacich72} of the designated individual. Let S be a scalar that adjusts the social value placed on status relative to constraint, as discussed below. The second terms captures the element of autonomy. Removing individual \textit{i} from the network, compute the eigenvectors of the remaining individuals and sum those that are connected to individual \textit{i}, taking the average eigenvector of neighbors \textit{j} as the constraint confronting node \textit{i} in the network. This is substantively the same mechanism as conducted in a jackknife resampling procedure, except that rather than forming a sampling distribution, we are measuring local constraint. Finally, subtract the status of the average neighbor connected to \textit{i} from \textit{i}’s scalar adjusted eigenvector centrality in the whole network.

If a node is an isolate, its eigenvector centrality should be set at 0. In the case of a network with no connections (such as a star graph), it can be accidentally calculated as 1.00 in programs. However, given that it is still a network with no connections, setting isolates at 0.00 is theoretically appropriate, considering that status is a reflection of connections. It would be inappropriate for an isolate to have the same status as the highest status node in a large network. Additionally, given the second step of subtraction, it is only theoretically meaningful if set as 0 in the case of distinction centrality.

\subsection{Scalar Adjustment}
First, a scalar is required because the relative value of status and autonomy is not inherently fixed in a social system. We might imagine situations where one desires more status at the expense of autonomy if the rewards to status are higher, or values autonomy more highly if the rewards to status are low. There cannot be an assumed unadjusted equivalence between the resulting eigenvectors of the two terms for this theoretical level.

Additionally, due to the hierarchical nature of networks, the average constraint in the network is greater than the average status. The high average constraint of their neighbors is often much larger than the impact of status. The result is that, at the level of the entire network, those who are deeply embedded among the most elite, without a basis of supporters, are very indistinct when considering the high constraints placed on them by their high-status alters. Thus, the calculated results without a scalar may fail to capture the appropriate value of status compared to the constraint.

While an arbitrary scalar is possible, it can yield erroneous results when applied to empirical networks. However, a scalar can be derived from the network itself due to the field positional interplay between status and constraint. Imagining two actors of equal status and personal constraint, one in an averagely more constrained system and the other in a less averagely constrained system, with average status otherwise equal. In such a case, the status of an actor in the more constrained system has a greater impact on their social distinction than in the less constrained system. With high constraint, status takes on greater importance in determining social distinction because it compels deference. In other words, it is the difference between a hierarchical system with and without constraints. Status with a weak constraint means a weaker compulsion of deference. The structure of constraint redounds into the impact of status on social distinction beyond its contribution.

With this social understanding of how constraint interacts with the transmission of status, an endogenous scalar can be derived from the network. The average constraint of actors within the network divided by the average status of actors in the network creates a scalar that reflects the relationship between status and constraint.

@@

Under conditions of high average constraint and low average status, the scalar is large, indicating the significant role of status in determining social distinction. Under conditions where the average constraint is low and the average status is high, the scalar is weaker, reflecting a system in which status is less important in determining social distinction. Under very large-scale conditions, the averages of status and constraint should trend towards 1, weakening the necessity of the scalar. However, under constrained conditions, particularly in elite networks with missing data on the social scaffolding that created the network elites, the scalar is more important for explaining action because the range of eigenvector network measures is otherwise artificially expanded due to the missing data.

Without a scalar, networks with any degree of hierarchy see the average constraint overwhelm the average status because many low-status nodes are tied to high-status nodes. The advantage of this scalar is to ensure that the average scalar status in the network equals the average constraint in the network. Thus, total status is equal to total constraint, and from there it becomes a matter of distribution among members. From there, however, it enables comparability between networks that would otherwise be impossible. With the understanding that we have, through the scalar, adjusted for a network where one's combination of status and constraint is better or worse, networks can be compared so long as the unobserved social context within the network remains stable.

@@

\textit{Model (b)} presents an idealized depiction of a structural fold \citep{vedres2010structural}. The resulting measurement scores are in Table 2. While node 1 maintains high eigenvector centrality, the eigenvector centrality of the other nodes increases, while their constraint decreases. It is actually the constraint of Node 1 that increases, reaching 1.00 because all its neighbors are maximally interconnected in their respective subgroups. This position of brokerage is more constrained than the star graph because the other actors are connected in subgroups. The actors in the subgroup are able to combine their interests to mitigate against the independence of Node 1 in the center. In an unadjusted measure of distinction, the different nodes become remarkably close in their distinction measure. With a scalar adjustment, the stratification increases, giving Node 1 a bit more privilege, but also demonstrating the extent to which Node 1's distinction is eroded. In a maximally connected graph, the distinction of each node would also go to 0.00, as there is no network differentiation among nodes.

\textit{Model (c)} presents a graph that is 2-connected in the absence of node 1. Each subgraph is one connection short of forming complete subgraphs in the absence of Node 1. Compared to the structural fold, connections 3-2 and 6-7 have been removed. Measurement results are in Table 3. Of the demonstrative networks here, this is the one that most approximates a three-tiered ordered hierarchy. The resulting scalar adjusted distinction measures well reflect this stratification, with Node 1 retaining its distinction, Nodes 4 and 5 failing in the middle position, and Nodes 2, 3, 6, and 7 falling to the bottom.

\textit{Model (d)} is the same 2-connected graph with the addition of a long tie connecting nodes 2 and 7. This long tie stratifies the network even more than in Model (c), bringing Nodes 2 and 7 quite close to 4 and 5. The formation of the long tie served to bolster the individual distinction of 2 and 7 more than the formation of the tie that would create the completely connected subgraphs as in Model (b). In fact, all nodes except 3 and 6 have higher distinction in such a case. This reflects the positional idea that distinction is a matter of stratification, and cohesion is at odds with creating stratification. While the status of Nodes 2 and 7 actually increases beyond that of Nodes 4 and 5 due to the long tie, the constraint also increases as they are now both further constrained by the influence of Node 1, their mutual neighbor.

\subsection{Implementation}
To conduct the analyses below, we use a custom function written for the \textit{R} statistical computing environment \citep{R} software.\footnote{The function can be found here \href{https://github.com/olizardo/Jilbert-Distinction-Centrality/blob/main/distinction.R}{https://github.com/olizardo/Jilbert-Distinction-Centrality/blob/main/distinction.R}}

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/star.png}
            \caption{Star Graph}
            \label{fig:star}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/sf.png}
            \caption{Structural Fold}
            \label{fig:sf}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/intra.png}
            \caption{2-Connected Subgraphs}
            \label{fig:intra}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/inter.png}
            \caption{2-Connected Subgraphs with Long Tie}
            \label{fig:inter}
    \end{subfigure}
    \caption{Distinction Centrality in toy graphs. Nodes with the highest distinction are shown in blue, nodes with the lowest distinction are shown in red, and nodes with distinction scores close to zero are shown in purple.}
    \label{fig:toys}
\end{figure}

@@

\section{Distinction as Search Algorithm}

\section{Distinction with Structural Holes vs. Structural Folds}

\section{Implications for Theories of Distinction}
\subsection{Long Ties and Opportunities for Distinction}
Given that networks are not static, but evolving, the incentives to maximize distinction within a network can inform the evolution of culture as a function of network reorganization. As cultural capital is an outcome of both status and uniqueness, the same calculus of optimizing network position can be extended.

@@

\subsection{Capital Conversion as Network Integration of Fields}
One of the problems with the relationship between multiple fields is how capital might be converted between fields. The ``rate'' between fields has always been a matter of confusion because it is difficult to quantify. Yet, it is an essential feature of an individual's navigation between the positional hierarchies of various fields. Seeking to convert a distinct position from one field to a better position in another is a critical strategy in the overall positional portfolio of agents.

@@

\subsection{Limitations on Positional Agency}
One problem with the above measure is that is creates the spectre of limitless positional agency: if a connection is mutually distinction increasing, actors should pursue it. Given that distinction increasing positional changes are possible, but go un-enacted, there must be constraints to distinction maximizing action. Theoretical considerations must be made to account for costs that constrict connection formation for distinction optimization.
